---
title: "Caret Example: Predicting SNI from NIHTB Measures"
output: html_notebook
---

### Variables of Interest

#### From NIH Toolbox (Social)
- NIHTB_Emotional_Support
- NIHTB_Friendship
- NIHTB_Loneliness
- NIHTB_Instrumetnal_Support
- NIHTB_Perceived_Hostility
- NIHTB_Perceived_Rejection

#### From Social Network Index
- SNI_Ex_People_Count
- SNI_Ex_Embed_Networks
- SNI_Ex_Network_Diversity

```{r}
# Import packages
library(splitstackshape)
library(plyr)
library(doBy)
library(gdata)
library(ggplot2)
library(Metrics)
library(data.table)
library(car)
library(mvtboost)
library(caret)
library(doMC)
library(lattice)

# Set working directory
setwd("/Users/jmt/GitHub/caret_demo")

# Read TSV data
X <- read.delim("SNI-NTB Results 2017-09-25.csv", sep=",", header=TRUE, fill=TRUE)

head(X)
```

### Plot matrix of pairwise scatterplots for all variables of interest
```{r fig.width = 12, fig.height = 12}
scatterplotMatrix(X, smooth = FALSE)
```

### Extract included NIHTB and SNI measures
```{r}
NTB <- X[, grepl("NIHTB", names(X))] * 1.0
SNI <- X[, grepl("SNI", names(X))] * 1.0

# Standardize variables [mean,sd] = [0,1]
NTBs <- scale(NTB)
SNIs <- scale(SNI)
```

### Compare tuned GBM and Stepwise Linear regression models using Caret
```{r}
registerDoMC(cores = 8)

# Collect data
social <- merge(NTBs, SNIs)
```

## Preprocessing checks

### Near-zero Variance Predictors
```{r}
nearZeroVar(NTBs, saveMetrics = TRUE)
```

### NTB Predictor Correlations
```{r}
corNTB <- cor(NTBs)
summary(corNTB[upper.tri(corNTB)])
```

### Number of highly correlated predictors
```{r}
findCorrelation(corNTB, cutoff = .75)
```

### Linear Combinations within Predictors
```{r}
findLinearCombos(NTBs)
```

### Setup 10 x 10-fold Cross Validation
```{r}
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 1)
```

### Set up univariate regression formulae
We'll only try to predict SNI people count from NTB in this example
```{r}
f1 = SNI_Ex_People_Count ~ . - SNI_Ex_Embed_Networks - SNI_Ex_Network_Diversity
f2 = SNI_Ex_Embed_Networks ~ . - SNI_Ex_People_Count - SNI_Ex_Network_Diversity
f3 = SNI_Ex_Network_Diversity ~ . - SNI_Ex_People_Count - SNI_Ex_Embed_Networks
```

### Tune the GBM Regressors for SNI Measures
```{r}
gbmGrid <-  expand.grid(interaction.depth = 2:5, 
                        n.trees = 10 * (1:5), 
                        shrinkage = c(0.01, 0.005),
                        n.minobsinnode = 10)

set.seed(1966)
```

### GBM : SNI People Count
```{r}
gbmFit.SNIcount <- train(f1,
                         data = social,
                         method = "gbm",
                         trControl = fitControl,
                         tuneGrid = gbmGrid,
                         verbose = FALSE)

gbmFit.SNIcount
```

```{r}
plot(gbmFit.SNIcount)
```

### Tuning Linear Regression with Backwards Selection
```{r}
set.seed(1966)
lrbackGrid <-  expand.grid(nvmax = 1:6)
lrbackFit.SNIcount <- train(f1,
                            data = social,
                            method = "leapBackward",
                            trControl = fitControl,
                            tuneGrid = lrbackGrid)

lrbackFit.SNIcount
```

```{r}
plot(lrbackFit.SNIcount)
```

### Multivariate gradient boosted tree regression
Predicting SNI measures from NIHTB Social Measures
```{r}
# Run MV Tree Regression with cross validation
mvtb.model <- mvtb(Y = SNIs, X = NTBs,
                   n.trees = 250, shrinkage = 0.05,
                   interaction.depth = 1, cv.folds = 10)

summary(mvtb.model)
```

### R2 for SNI predictions from NTB
```{r}
SNIs.hat <- predict(mvtb.model, newdata=NTBs)
SNIs.R2 <- diag(var(SNIs.hat)/var(SNIs))
SNIs.R2
```

### RMSE for SNI prediction from NTB
```{r}
SNIs.RMSE <- rmse(SNIs.hat, SNIs)
SNIs.RMSE
```

### Relative influences of NIHTB Social on SNI
```{r}
par(mar=c(15,12,2,2))
relinfl <- t(mvtb.ri(mvtb.model, relative="tot"))
mvtb.heat(relinfl, clust.method="ward.D")
```

### Outcome covariance explained
```{r}
covex <- mvtb.covex(mvtb.model, Y=SNIs, X=NTBs)
par(mar=c(12,25,2,2))
mvtb.heat(covex, clust.method="ward.D")
```

### Training and Cross-validation errors
```{r}
Trees <- 1:length(mvtb.model$trainerr)
errs <- data.frame(Trees, mvtb.model$trainerr, mvtb.model$cv.err)
errs <- reshape2::melt(errs,id="Trees")
vp <- ggplot(errs, aes(x=Trees, y=value, color=variable))
vp <- vp + geom_line()
vp
```