---
title: "Caret Preprocessing Examples"
output: html_notebook
---

Adapted from https://topepo.github.io/caret/pre-processing.html

```{r}
library(caret)
library(ggplot2)
```


## Identifying Linear Dependencies between Predictors
Set up a 6 x 6 matrix where the first column is the sum of columns 2 and 3, and also of columns 4, 5 and 6
```{r}
ltfrDesign <- matrix(0, nrow=6, ncol=6)
ltfrDesign[,1] <- c(1, 1, 1, 1, 1, 1)
ltfrDesign[,2] <- c(1, 1, 1, 0, 0, 0)
ltfrDesign[,3] <- c(0, 0, 0, 1, 1, 1)
ltfrDesign[,4] <- c(1, 0, 0, 1, 0, 0)
ltfrDesign[,5] <- c(0, 1, 0, 0, 1, 0)
ltfrDesign[,6] <- c(0, 0, 1, 0, 0, 1)
ltfrDesign
```

Now find linear combinations of columns and indicate which columns can be removed. The function ```findLinearCombos``` uses the QR decomposition of a matrix to enumerate sets of linear combinations between predictors (columns)

```{r}
comboInfo <- findLinearCombos(ltfrDesign)
comboInfo
ltfrDesign[, -comboInfo$remove]
```

## Identifying Zero and Near-zero Variance Predictors

This example uses the Multidrug Resistance Reversal (MDRR) dataset bundled with R. Each of the 342 tested compounds has 528 descriptors of its molecular structure.

```{r}
data(mdrr)
```

Here's a reasonably well-blanced structral descriptor (nR10, the number of 10-membered rings in the compound)
```{r}
library(ggplot2)
qplot(mdrrDescr$nR10,
      geom="histogram",
      bins = 5,
      binwidth = 0.5,
      xlab = "nR10")
```

One of the structural descriptors (nR11, the number of 11-membered rings) has only a few unique numerical values (0, 1, 2) that are highly unbalanced.

```{r}
qplot(mdrrDescr$nR11,
      geom="histogram",
      bins = 5,
      binwidth = 0.5,
      xlab = "nR11")
```
The concern here is that these predictors may become zero-variance predictors when the data are split into cross-validation/bootstrap sub-samples or that a few samples may have an undue influence on the model.

```{r}
nzv <- nearZeroVar(mdrrDescr, saveMetrics = TRUE)
nzv[40:50,]
```
```freqRatio``` is the ratio of the frequencies of the most common and second most common value. Values >> 1 indicate problem predictors. ```percentUnique``` is the percentage of the all values that are unique. This should be close to zero for well-behaved predictors..

nearZeroVar returns a boolean vector which can be used to remove problematic near zero variance predictors (45 in this case, including nR11)

```{r}
nzv_new <- nearZeroVar(mdrrDescr)
filteredDescr <- mdrrDescr[, -nzv_new]
dim(mdrrDescr)
dim(filteredDescr)
```

